Proyecto Final para la obtención del título: Ingeniero en Informática \tabularnewline \large Future Virtual Particle Method for Pedestrian Navigation

### 2015-06-09 doc/informe/informe.tex
elimino varios archivos tex con cosas viejas. Actualizo imagenes. Actualizo informe.

\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{graphicx}

@@

%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

\usepackage{babel}

@@

\usepackage{babel}
\begin{document}

\title{Future Virtual Particle Method for Pedestrian Navigation}

\author{
  Castiglione, Gonzalo\\
  \texttt{gcastigl@gmail.com}
  \and
  Marseillan, Agustin\\
  \texttt{agustinmarseillan@gmail.com}
  \and
  Parisi, Daniel\\
  \texttt{danielparisi@xxxxx.com}
}
\date{}

@@

\vspace{10cm}
\begin{center}
    \includegraphics[scale=0.08]{pics/ITBA}
    \par
\end{center}

@@

\begin{abstract}
Pedestrians are modeled as passive objects affected by repulsive and
attractive forces in the Social Force Model (SFM). This force governed
model leads to unnatural behaviors that don't resemble reality. This
paper presents an avoidance collision method based in pedestrian self
governed decisions, by calculating the position of every pedestrian
in the future, a given pedestrian can adjust his velocity vector to
avoid collisions instead of a repulsive force.

\textbf{keywords:} pedestrian, collision avoidance, future virtual
particle, force model.
\end{abstract}

@@

Navigation of biological, sinthetic or virtual agents is a relevant
problem in several fields such as pedestrian dynamics, moving robots
and animation of characters for videogames and motion pictures.

@@

\begin{equation}
\vec{e_{i}}=\frac{\vec{x}_{i}^{k}-\vec{x}_{i}(t)}{||\vec{x}_{i}^{k}-\vec{x}_{i}(t)||}\label{eq:desired-direction}
\end{equation}

@@

$x_{i}^{k}$ is the closest point from the goal (represented as an
area) to pedestrian $i$. \\

Figure \ref{fig:driving-force} shows a pedestrian moving with $\vec{v}$
velocity but adjusting its trajectory towards \textbf{$X$}.

\begin{figure}[h]
\begin{centering}
\includegraphics[scale=0.4]{pics/sfm/drivingforce} 
\par\end{centering}

\caption{\label{fig:driving-force}Driving force}
\end{figure}

@@

Figure \ref{fig:Social-Force} shows \ref{eq:social-force} graphically.

\begin{figure}[h]
\begin{centering}
\includegraphics[scale=0.4]{pics/sfm/sf} 
\par\end{centering}

\caption{Social Force\label{fig:Social-Force}}
\end{figure}

The third term is known as the ``Granular force''. It's calculated
as follows:

\begin{equation}
\vec{F}_{G_{i}}=\sum_{j=1,j\ne i}^{N_{P}}[-\epsilon_{ij}k_{n}\vec{e}_{ij}^{n}+v_{ij}^{t}\epsilon_{ij}k_{t}\vec{e}_{ij}^{t}]\, g(\epsilon_{ij})\label{eq:granular-force}
\end{equation}

@@

Figure \ref{fig:colliding-pedestrians} shows \ref{eq:granular-force}
graphically.

\begin{figure}[h]
\begin{centering}
\includegraphics[scale=0.4]{pics/sfm/ganular} 
\par\end{centering}

\caption{Colliding pedestrians\label{fig:colliding-pedestrians}}
\end{figure}

@@

In the FVPM, the social force in equation \ref{eq:driving-force}
is eliminated and replaced with a dynamic driving force towards a
short term objective, which is equal to equation \ref{eq:desired-direction}
replacing $\vec{x}_{i}^{k}$ with $\vec{x_{i'}}$, where $\vec{x_{i'}}$
is an estimation of the location of pedestrian $i$ in the near future.
$i'$ will be called Future Virtual Particle.

AGREGADO POR PARISI: LLEGAMOS A POSTULAR ESTE MODELO DESPUES DE VARIOS CICLOS DE ITERACIONES Y NUMEROSAS PRUEBAS DE DE VARIANTES HASTA CONVERGER AL MODELO PRESENTADO QUE MODELA ADECUADAMENTE VARIAS SITUACIONES

\section{The Model}

@@

The heuristics that govern the motion of a pedestrian are the same
as Helbin's \cite{key-helb1995,key-helb2000}:

@@

    Represents the personal space of a pedetrian. The value of the radio
    is generated randomly for each pedestrian. The range of values is
    distributed uniformly in $[0.25,0.29]$ $[cm]$ between pedestrians.

@@

\begin{figure}[h]
    \centering{}
    \includegraphics[scale=0.4]{pics/pedestrian-top} 
    \caption{\label{fig:pedestrian-geometry}Illustration of pedestrian $i$ and its FVP ($i'$).}
\end{figure}

The naming used is shown in illustration \ref{fig:pedestrian-vectors}.

\begin{figure}[h]
    \centering{}
    \includegraphics[scale=0.4]{pics/geometry}
    \caption{\label{fig:pedestrian-vectors}Vectorial definitions}
\end{figure}

@@

To model this situations, two types of forces act over the FVP: 
\begin{itemize}
\item Internal force \\

This force will guide the pedestrian on the shortest path $\mathbf{x_{o}x_{i}}$.
This force is modeled using two springs:

The first spring starts on $x_{i}$, ends on $x_{i'}$ and has a stationary
distance of $RD$ with a spring constant $K_{s}(\theta)$ where $\theta$
is the angle between $\mathbf{u}$ and $\mathbf{r}$. With this spring
the FVP will tend to always be at $RD$ distance from the pedestrian.
Because a pedestrian always tries to reach his goal in the shortest
possible path (hypothesis 1), if it has to take a big detour of his
ideal path, it will try to reduce his velocity drastically in order
to avoid making a long travel. To recreate this, the spring constant
has to be dependant of the deviation angle. 
\[
K_{s}(\theta)=\frac{K_{2}}{\theta}
\]

The second spring starts on $x_{i'}$ and ends on $x_{i}+RD\frac{\vec{x_{i}x_{o}}}{\mid\vec{x_{i}x_{o}}\mid}$
with a spring constant $K_{1}$. This spring represents the pedestrian's
desire to reach his target. A greater spring constant will force a
straighter path. In order to avoid an oscillatory movement, a damping
$\gamma$ was added to the spring.\\

@@

\begin{equation}
    F_{internal}(i)=F_{S1}(i)+F_{S2}(i)
\end{equation}

\[
    F_{S1}(i)=K_{1}\cdot(|\vec{r}|-RD)\cdot\hat{r}
\]
\[
    F_{S2}(i)=\, K_{2}(\theta_{i})\cdot(\vec{x_{i'}}-(\vec{x_{i}}+\hat{x_{i}o}\cdot RD))
\]

\begin{figure}[h]
    \centering{}\includegraphics[scale=0.35]{pics/pedestrian-internal-forces}
    \caption{Pedestrian internal forces}
\end{figure}

\item Avoidance forces \\
 This force will produce avoidance movements when obstacles are detected
on the path. The magnitude of this force is calculated using two factors:
\\
 The first factor is calculated for each of the pedestrian $j$ ($j\neq i$)
who is in the range of sight of pedestrian $i$. This restriction
is verified using the following condition:

\[
    \mathbf{r_{ii'}}\bullet\mathbf{r_{ij}}<0
\]

This condition represents the fact that pedestrians make decisions
based only on obstacles in his range of vision. The formula for the
external force that affects the FVP $i'$ is

\[
    F_{ext}(i')=FF(i')+FP(i')+FW(i')
\]
where 
\[
    FF_{(}i')=\sum_{j}(\alpha_{ff}e^{-s/\beta_{ff}})
\]
\[
    FP_{(}i')=\sum_{j}(\alpha_{fp}e^{-t/\beta_{fp}})
\]
\[
    FW_{(}i')=\sum_{w}(\alpha_{fw}e^{-dist(w,i')/\beta_{fw}})
\]

where $\alpha_{x}$ and $\beta_{x}$ are constants.

The first term of the sum acts as a repulsion force between $i'$
and $j'$, resulting in the avoidance of a future collision. The second
term uses this same principle but calculates the repulsion force between
$j'$ and $i$. The third term adds the repulsion against walls, using
the closest point between $i'$ and the wall.

Figure \ref{fig:External-forces-affecting} shows the external forces
that a pedestrian $i$ suffers because of another pedestrian ($j$)
and also the direction in which $i$ desires to move.

\begin{figure}[h]
    \centering{}
    \includegraphics[scale=0.4]{pics/pedestrian-top-forces}
    \caption{\label{fig:External-forces-affecting}External forces affecting future$i'$}
\end{figure}

@@

After this, the final $F_{i'}$ is computed by adding all the terms.
$F_{i'}=F_{ext}(i)+F_{i\, s1}+F_{i\, s2}$.\\

To avoid high symetry situations, a low noise $P=10\%$ is added to
$F$. There are two ways to apply this noise: 
\begin{itemize}

\item Radial noise:

\begin{itemize}
\item A value $p$ is taken randomly from a uniform distribution $[-P,\, P]$
and calculate: $FL_{i'}=F_{i'}*p$ 
\item Angular noise:

    \begin{itemize}
        \item A value $sgn=\{-1,\,1\}$ is taken randomly from a uniform distribution
        and a value $p$ from $[-P,\, P]$. Then \\
        \[
            M=\left(\begin{array}{cc}
            \cos(\pi*sgn) & -\sin(\pi*sgn)\\
            \sin(\pi*sgn) & \cos(\pi*sgn)
            \end{array}\right)
        \]
        \[
            FA_{i'}=F_{i'}*p*M
        \]
    \end{itemize}
\end{itemize}

At last, we find $F'_{i'}=F_{i'}+FL_{i'}+FA_{i'}$ and apply movement
equations.
\end{itemize}
\item Dynamic of the pedestrian

The pedestrian always wants to move in the direction his FVP is pointing
and its magnitude is defined as $F_{d}$ or driving force:

\[
    \vec{F}_{desire_{i}}=m_{i}\frac{\frac{|\vec{ii'}|}{|\vec{r}_{max}|}\vec{e_{i}}-\vec{v}_{i}}{\tau}
\]

Where $\tau=0.5$

\textit{{*}It is important to note that the only deviation from the
SFM in this equation is the $k\vec{e}_{i}$ term. As described above,
this model focuses on improving the SFM through dynamically adjusting
the desired velocity.}

@@

The pedestrian movement is calculated in four steps: 
\begin{enumerate}
    \item Calculate forces for each FVP. 
    \item Update positions for each FVP. 
    \item Calculate forces for each pedestrian. 
    \item Update positions for each pedestrian. 
\end{enumerate}

@@

\subsection{Metrics}

The results where compared to the SFM model \cite{key-helb2000}.
The test scenarios where crossing and hallway for they present the
main types of symmetry (90 degrees and 180 degrees respectively). \\

Hallway scenario: A $15[m]$ by $4[m]$ hallway. Pedestrians are generated
from each end at a pace of 1 pedestrian per second with the other
end as target. \\
 This scenario is shown in figure \ref{fig:hallway-scenario}.

\begin{figure}[h]
    \centering{}
    \includegraphics[scale=0.35]{pics/scenarios/hallway}
    \caption{\label{fig:hallway-scenario}Hallway Scenario}
\end{figure}

Crossing scenario: Two hallways of $25[m]$ lenght by $5[m]$ width
put toghether in the center forming a cross. Pedestrians are generated
from the top, targeting the bottom, and from the right end, targeting
the left end, at a pace of 1 pedestrian per second. \\
 This scenario is shown in figure \ref{fig:cross-scenario}.

\begin{figure}[h]
    \centering{}
    \includegraphics[scale=0.3]{pics/scenarios/cross}
    \caption{\label{fig:cross-scenario}Cross Scenario}
\end{figure}

The metrics used to validate and verify the model against the defined
two previous scenarios are defined as follows:
\begin{enumerate}
\item Number of collisions: The equation for this metric is defined as:
\\

\begin{eqnarray*}
    \sum\#\{(i,j)\,/\, id_{i}<id_{j}\wedge\\
    x_{i,t_{n}}-x_{j,t_{n}}<R_{i}+R_{j}\wedge\\
    x_{i,t_{n-1}}-x_{j,t_{n-1}}>R_{i}+R_{j}\}\,\,\forall t_{n}
\end{eqnarray*}

\item Amount of collitions per instant: The equation for this metric is
defined as: \\

\begin{eqnarray*}
    \sum\#\{(i,j)\,/\, id_{i}<id_{j}\wedge\\
    x_{i,t_{n}}-x_{j,t_{n}}<R_{i}+R_{j}\}\,\,\forall t_{n}
\end{eqnarray*}

\item Average walking speed: The equation for this metric is defined as:
\\

\begin{eqnarray*}
    \sum v_{i,t_{n}}/\#\{i_{t_{n}}\}\,\,\forall t_{n}
\end{eqnarray*}

\item Average travel time: The average time that a pedestrian needed until
it reached the goal. The equation for this metric is defined as: \\

\begin{eqnarray*}
\sum\#\{i_{t_{n}}\}*\triangle t\,\,\forall t_{n}/\#\{i\}
\end{eqnarray*}

\item Average travel distance: The average distance that a pedestrian traveled
until it reached the goal. The equation for this metric is defined
as: \\

\begin{eqnarray*}
\sum\#\{v_{i,t_{n}}\}*\triangle t\,\,\forall t_{n}/\#\{i\}
\end{eqnarray*}

\item Average turn angle: The average angle turned by a pedestrian until
it reached the goal.\\

\begin{eqnarray*}
    \sum\arccos(v_{t_{n}}\bullet v_{t_{n-1}}/|v_{t_{n}}\bullet v_{t_{n-1}}|)\,\,\forall t_{n}/\#\{i\}
\end{eqnarray*}

\end{enumerate}
Escape room: A $20x20[m]$ room with a single $1.2[m]$ exit door
centered at the bottom. At the beginning, $100$ pedestrians are distributed
across the room and target the exit. When the first pedestrians exits
the room, a counter starts, and when the last one croses, it stops.
This way we can measure the flow of pedestrians leaving a room, which
is known to be between $0.69$ and $1$ for pedestrians without emergencies
\cite{key-pedEvDyn2012}. \\
 This scenario is shown in figure \ref{fig:room}.

\begin{figure}[h]
    \centering{}
    \includegraphics[scale=0.3]{pics/scenarios/room}
    \caption{\label{fig:room}Room Scenario}
\end{figure}

The metrics used where: 
\begin{enumerate}
    \item Escape flow: 
    \begin{eqnarray*}
    \sum\arccos(v_{t_{n}}\bullet v_{t_{n-1}}/|v_{t_{n}}\bullet v_{t_{n-1}}|)\,\,\forall t_{n}/\#\{i\}
\end{eqnarray*}

@@

@@

To calibrate the model, runs varying parameters were made. A wide
spectrum of values was covered, testing every combination of every
possible one. After seeing clear preferences towards certain values,
the values were refined within that scope. After numerous iterations
of this process, the values that best suit these metrics are:

@@

    \begin{itemize}
        \item $K_{1}=100$ {[}N/m{]}, 
        \item $\gamma=10$ 
        \item $K_{2}=150$ {[}N/m{]} 
    \end{itemize}
    \item FVP-FVP interaction

    \begin{itemize}
        \item $\alpha=1000$ 
        \item $\beta=[0.4,\,0.5]$ (Uniform distribution) 
        \end{itemize}
    \item Pedestrian-FVP interaction

@@

Metric results for crossing scenario:

\begin{tabular}{|c|c|c|}
    \hline 
    Metric  & {\scriptsize{FVPM}} & {\scriptsize{SFM ($\alpha=2000,\,\beta=0.08$)}}\tabularnewline
    \hline 
    \hline 
    $1$  & {\scriptsize{$1.800\pm20.748$}} & {\scriptsize{$5.333\pm2.625$}}\tabularnewline
    \hline 
    $2$  & {\scriptsize{$34.600\pm8.333$}} & {\scriptsize{$12.333\pm8.340$}}\tabularnewline
    \hline 
    $3$  & {\scriptsize{$1.024\pm0.016$ $[m/s]$}} & {\scriptsize{$1.052\pm0.003$ $[m/s]$}}\tabularnewline
    \hline 
    $4$  & {\scriptsize{$1.910\pm0.022$ $[s]$}} & {\scriptsize{$1.876\pm0.003$ $[s]$}}\tabularnewline
    \hline 
    $5$  & {\scriptsize{$2.392\pm0.005$ $[m]$}} & {\scriptsize{$2.391\pm0.002$ $[m]$}}\tabularnewline
    \hline 
    $6$  & {\scriptsize{$91.062\pm12.122$ $[rad]$}} & {\scriptsize{$106.185\pm13.778$ $[rad]$}}\tabularnewline
    \hline 
\end{tabular}

\begin{figure}[h]
    \begin{centering}
    \includegraphics[width=6cm]{pics/program/crossing-no-future} 
    \par\end{centering}
    \caption{\label{fig:crossing-no-future}Crossing visualization}
\end{figure}

\begin{figure}[h]
    \centering{} \includegraphics[width=6cm]{pics/program/crossing-future}
    \caption{\label{fig:crossing-future}Crossing visualization with visible future}
\end{figure}

Metric results for hallway scenario:

>>>>

\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}

\usepackage{float}
\usepackage{babel}

\usepackage{array}
\usepackage{multirow}

\usepackage[font=small,labelfont=bf,labelsep=space]{caption}
\captionsetup{
  figurename = Figure,
  tablename = Table
}

@@

\providecommand{\tabularnewline}{\\}

@@

\begin{document}

\title{
    Proyecto Final para la obtención del título: Ingeniero en Informática
    \tabularnewline
    \large Future Virtual Particle Method for Pedestrian Navigation
}

\author{
  Castiglione, Gonzalo\\
  \texttt{gcastigl@gmail.com}
  \and
  Marseillan, Agustin\\
  \texttt{agustinmarseillan@gmail.com}
  \and
  Dr. Parisi, Daniel Ricardo \\
  \texttt{dparisi@itba.edu.ar}
}
\date{}

@@

\vspace{8cm}
\begin{center}
    \includegraphics[scale=0.1]{pics/ITBA}
    \par
\end{center}

@@

\tableofcontents{}

\begin{abstract}
% Pedestrians are modeled as objects affected by repulsive and attractive forces in the Social Force Model (SFM). 
This paper presents an avoidance collision method based in pedestrian self
governed decisions, by calculating the position of every pedestrian
in the future, a given pedestrian can adjust his velocity vector to
avoid collisions instead of being affected by a repulsive force. \\
The model was tested using three different scenarios and compared 
against the Social Force Model obtaining more natural navigation,
reduced number of collisions, increased average
speed and reduced average deviation angle of navigating pedestrians.
\textbf{keywords:} pedestrian, collision avoidance, future virtual
particle, force model.
\end{abstract}

@@

    Navigation of biological, synthetic or virtual agents is a relevant
    problem in several fields such as pedestrian dynamics, moving robots
    and animation of characters for video games and motion pictures.

@@

    \begin{equation}
    \vec{e_{i}}=\frac{\vec{x}_{i0}-\vec{x}_{i}(t)}{||\vec{x}_{i0}-\vec{x}_{i}(t)||}\label{eq:desired-direction}
    \end{equation}

@@

    $x_{i0}$ is the closest point from the goal (represented as an
    area) to pedestrian $i$. \\

    Fig. \ref{fig:driving-force} shows a pedestrian moving with $\vec{v}$
    velocity but adjusting its trajectory towards \textbf{$X$}.

    \begin{figure}[h]
        \begin{centering}
        \includegraphics[scale=0.4]{pics/sfm/drivingforce} 
        \par
        \end{centering}
        \caption{\label{fig:driving-force} Driving force}
    \end{figure}

@@

    Fig. \ref{fig:Social-Force} shows equation \ref{eq:social-force} graphically.

    \begin{figure}[h]
        \begin{centering}
        \includegraphics[scale=0.4]{pics/sfm/sf} \par
        \end{centering}
        \caption{Social Force\label{fig:Social-Force}}
    \end{figure}

    The third term is known as the ``Contact force''. It's calculated
    as follows:

    \begin{equation}
        \vec{F}_{G_{i}} = \sum_{j=1,j\ne i}^{N_{P}} [-\epsilon_{ij}k_{n}\vec{e}_{ij}^{n} + v_{ij}^{t} \epsilon_{ij} k_{t} \vec{e}_{ij}^{t}] \, g(\epsilon_{ij})
        \label{eq:granular-force}
    \end{equation}

@@

    Fig. \ref{fig:colliding-pedestrians} shows \ref{eq:granular-force} graphically.

    \begin{figure}[h]
        \begin{centering}
        \includegraphics[scale=0.4]{pics/sfm/ganular} \par
        \end{centering}
        \caption{Colliding pedestrians\label{fig:colliding-pedestrians}}
    \end{figure}

@@

    In the FVPM, the social force in equation \ref{eq:driving-force}
    is eliminated and replaced with a dynamic driving force towards a
    short term objective, which is equal to equation \ref{eq:desired-direction}
    replacing $\vec{x}_{i}^{0}$ with $\vec{x_{i'}}$, where $\vec{x_{i'}}$
    is an estimation of the location of pedestrian $i$ in the near future.
    $i'$ will be called Future Virtual Particle. \\

    The presented work demanded a deep understanding of the problem in question, for which a 
    great study of related work was done. After that, we proposed various different models which
    were validated and reformed until it converged to this model, which validate against real
    world scenarios and metrics.
\section{The Model}

@@

    The basic concepts below the model of pedestrian movement are the same
    as Helbin's \cite{key-helb1995,key-helb2000}:

@@

        Represents the personal space of a pedestrian. The value of the radio
        is generated randomly for each pedestrian. The range of values is
        distributed uniformly in $[0.25,0.29]$ $[cm]$ between pedestrians.

@@

    \begin{figure}[h]
        \centering{}
        \includegraphics[scale=0.4]{pics/pedestrian-top} 
        \caption{\label{fig:pedestrian-geometry} Pedestrian $i$ and its FVP ($i'$).}
    \end{figure}

    The naming convention used for all vectors related to a pedestrian is shown in 
    fig. \ref{fig:pedestrian-vectors}.

    \begin{figure}[h]
        \centering{}
        \includegraphics[scale=0.4]{pics/geometry}
        \caption{\label{fig:pedestrian-vectors} Vectors definitions}
    \end{figure}

@@

        To model this situations, two types of forces act over the FVP:

        \begin{itemize}
        \item Internal force \\

@@

            \begin{equation}
                F_{i'}^{internal} = F_{i'}^{S1} + F_{i'}^{S2}
            \end{equation}
            where
            \[
                F_{i'}^{S1} = K_{1}(\theta_{i}) * (|\vec{r}| - RD) \hat{r}
            \]
            \[
                F_{i'}^{S2} = K_{2} * (\vec{x_{i'}} - (\vec{x_{i}} + \hat{x_{io}} RD))
            \]

            This force aligns the future on the path $\mathbf{x_{io}}$ and is modeled using 
            two springs $S_1$ and $S_2$:\\

            $S_1$ starts on $\mathbf{x_{i}}$, ends on $\mathbf{x_{i'}}$ and 
            has a stationary distance of $RD$ with a spring constant $K_{1}(\theta)$ 
            where $\theta$ is the angle between $\mathbf{u}$ and $\mathbf{r}$. With this 
            spring the FVP will tend to always be at $RD$ distance from the pedestrian.
            Because a pedestrian always tries to reach his goal in the shortest
            possible path (hypothesis 1), if it has to take a big detour of his
            ideal path, it will try to reduce his velocity drastically in order
            to avoid making a long travel. To recreate this, the spring constant
            has to be dependant of the deviation angle.
            \[
                K_{1}(\theta) = \frac{K_{1c}}{\theta}
            \]
            where $K_{1c}$ is a constant value.

            $S_2$ starts on $\mathbf{x_{i'}}$ and ends on $\mathbf{x_{i}} + RD\mathbf{{\hat{x}_{io}}}$
            with a spring constant $K_{2c}$. Setting a greater $K_{2c}$ spring constant 
            will force a straighter path but may result in more collisions. In order to 
            avoid an oscillatory movement, a damping $\gamma$ is added to the spring $S_2$. \\
            Fig. \ref{fig:internal-forces} shows the internal forces that a FVP $i'$ suffers 
            because it has to be aligned with the pedestrian's position $\mathbf{x_{i}}$ and the long 
            term objective $\mathbf{x_{o}}$ but also at distance $RD$ from $\mathbf{x_{i}}$.

            \begin{figure}[H]
                \centering{}
                \includegraphics[scale=0.35]{pics/pedestrian-internal-forces}
                \caption{\label{fig:internal-forces} Internal forces}
            \end{figure}

        \item External forces \\

            This force will produce avoidance movements when obstacles are detected
            on the path. The equation for the external force that affects the FVP $i'$ is
            \begin{equation}
                F_{i'}^{external} = FF_{i'} + FP_{i'} + FW_{i'} \label{eq:externalforce}
            \end{equation}
            where 
            \[
                FF_{i'} = \sum_{j}^{N}(\alpha_{ff}e^{-s/\beta_{ff}})
            \]
            \[
                FP_{i'} = \sum_{j}^{N}(\alpha_{fp}e^{-t/\beta_{fp}})
            \]
            \[
                FW_{i'} = \sum_{k}^{walls}(\alpha_{fw}e^{-\zeta_{k} /\beta_{fw}})
            \]
            having $\zeta_{k} = |x_{i'} - x_{k}|$; $x_{k}$ being the closest point from wall 
            $k$ to $x_{i'}$. \\
            $\alpha$ and $\beta$ are constants. \\
            The term $FF_{i'}$ in equation \ref{eq:externalforce} acts as a repulsion force 
            between $i'$ and $j'$, resulting in the avoidance of a future collision. The term 
            $FP_{i'}$ in equation \ref{eq:externalforce} uses this same principle but 
            calculates the repulsion  force between $i'$ and $j$. The term $FW_{i'}$ in equation 
            \ref{eq:externalforce} adds the repulsion against walls,  using the closest point 
            between $i'$ and the wall. \\
            This force is only applied to each the pedestrian $j$ ($j\neq i$)
            who is in the range of sight of pedestrian $i$. The restriction
            is verified using the following condition:
            \[
                \mathbf{r_{ii'}}\bullet\mathbf{r_{ij}} > 0
            \]
            The condition represents the fact that pedestrians make decisions
            based only on obstacles in his range of vision. \\
            Fig. \ref{fig:external-forces} shows the external forces
            that a FVP $i'$ suffers because of another pedestrian ($j$)
            and also the direction in which $i$ desires to move.

            \begin{figure}[H]
                \centering{}
                \includegraphics[scale=0.35]{pics/pedestrian-top-forces}
                \caption{\label{fig:external-forces} External forces}
            \end{figure}

@@

        After this, $F_{i'}$ is computed by adding all the terms.

        \begin{equation}
            F_{i'} = F_{i'}^{external} + F_{i'}^{internal} \label{eq:future-force}
        \end{equation}

        To avoid high symmetry situations, a small noise is added to $F$. 
        This noise is calculated by taking a random value $s$ with a random distribution from $\{-1, 1\}$
        and applying:
        \[
            M=\left(\begin{array}{cc}
            0 & -s\\
            s & 0
            \end{array}\right)
        \]
        in
        \[
            F'_{i'} = F_{i'} + F_{i'} * M
        \]

        Finally, movement equations are applied using $F'_{i'}$.

    \item Dynamic of the pedestrian

        The driving force $\vec{F}_{d_{i}}$ makes the pedestrian tend to move towards his FVP at velocity $v_{di}$
        \[
            \vec{F}_{d_{i}} = m_{i}\frac{|v_{di}|\frac{\vec{r}}{RD} - \vec{v}_{i}}{\tau}
        \]
        where $\tau=0.5$\\

        It is important to note that the only deviation from the
        SFM in this equation is the $k\vec{e}_{i}$ term. As described above,
        this model focuses on improving the SFM through dynamically adjusting
        the desired velocity.

@@

    The pedestrian movement is calculated in four steps: 
    \begin{enumerate}
        \item Calculate forces for each FVP. 
        \item Calculate forces for each pedestrian. 
        \item Update positions for each FVP. 
        \item Update positions for each pedestrian. 
    \end{enumerate}

@@

    \subsection{Scenarios}

        The test scenarios where crossing and hallway for they present the
        main types of symmetry (90 degrees and 180 degrees respectively). \\

        Hallway scenario: A $15[m]$ by $4[m]$ hallway. Pedestrians are generated
        from each end at a pace of 1 pedestrian per second with the other
        end as target. \\
         This scenario is shown in fig. \ref{fig:hallway-scenario}.

        \begin{figure}[H]
            \centering{}
            \includegraphics[scale=0.35]{pics/scenarios/hallway}
            \caption{\label{fig:hallway-scenario}Hallway Scenario}
        \end{figure}

        Crossing scenario: Two hallways of $25[m]$ length by $5[m]$ width
        put together in the center forming a cross. Pedestrians are generated
        from the top, targeting the bottom, and from the right end, targeting
        the left end, at a pace of 1 pedestrian per second. \\
         This scenario is shown in fig. \ref{fig:cross-scenario}.

        \begin{figure}[H]
            \centering{}
            \includegraphics[scale=0.3]{pics/scenarios/cross}
            \caption{\label{fig:cross-scenario} Cross Scenario}
        \end{figure}

        Room evacuation: A $20x20[m]$ room with a single exit door
        centered at the bottom. At the beginning, a fixed number of pedestrians are 
        distributed across the room. Each of them targeting the exit door. When the 
        first pedestrian exits the room, a counter starts, and when the last one 
        crosses, it stops. This way we can measure the flow of pedestrians leaving a 
        room, which is known to be between $1$ and $4$ [p/m/s] for pedestrians 
        without emergencies \cite{key-pari2009}. \\
        The door size will be tested using the following measures: $1.2[m]$, $1.5[m]$ and $1.8[m]$. \\
        This scenario is shown in fig. \ref{fig:room}.
        \begin{figure}[H]
            \centering{}
            \includegraphics[scale=0.3]{pics/scenarios/room}
            \caption{\label{fig:room} Room Scenario}
        \end{figure}

    \subsection{Metrics}
        The metrics used to validate and verify the model against the hallway and 
        crossing scenarios are defined as follows:

        \begin{enumerate}
            \item Number of collisions:

            \begin{eqnarray*}
                \sum_t\sum_i\sum_j collides(i,j)/2
            \end{eqnarray*}

            \begin{eqnarray*}
            collides(i,j,t) = touching(i,j,t) (1 - touching(i,j,t-1))\\
            \end{eqnarray*}

            \begin{eqnarray*}
            touching(i,j,t) = 
            \left \{
                \begin{tabular}{cc}
                $0$ & if $t<=0$\\
                $1$ & if $x_i-x_j<R_i+R_j$\\
                $0$ & if not
                \end{tabular}
            \right .
            \end{eqnarray*}

            \item Amount of collisions per instant: The equation for this metric is
            defined as:

            \begin{eqnarray*}
                \sum_t\sum_i\sum_j touching(i,j)/2
                %\sum\#\{(i,j)\,/\, id_{i}<id_{j}\wedge\\
                %x_{i,t_{n}}-x_{j,t_{n}}<R_{i}+R_{j}\}\,\,\forall t_{n}
            \end{eqnarray*}

            \item Average walking speed:

            \begin{eqnarray*}
                \sum_{t=0}^T((\sum_{i=0}^N v_i)/N)/T
            \end{eqnarray*}

            \item Average travel time: The average time that a pedestrian takes to
            reach the goal. This is calculated only for pedestrians who reach the goal
            before the simulation ends.

            \item Average travel distance: The average distance that pedestrians traveled
            until it reached the goal. This is calculated only for pedestrians who reach the goal
            before the simulation ends.

            \item Average turn angle: The average angle turned by a pedestrian until
            it reached the goal. This is calculated only for pedestrians who reach the goal
            before the simulation ends.\\
            where the turn angle is:

            \begin{eqnarray*}
                \arccos(v_{t_{n}}\bullet v_{t_{n-1}}/|v_{t_{n}}\bullet v_{t_{n-1}}|)
            \end{eqnarray*}

@@

        The metric used for the room evacuation scenario is:

        \begin{enumerate}
            \item Escape flow: The amount of pedestrians coming out of the room by second.
            \begin{eqnarray*}
                \frac{N}{\Delta T}
            \end{eqnarray*}

            where $\Delta T$ is the time when the last pedestrian exits minus
            the time when the first pedestrian exits.

        All results were compared to the SFM model \cite{key-helb2000}.

@@

        To calibrate the model, runs with varying parameters were made.
        % MARSE TRADUCIR: Los parametros fueron buscados de manera exhaustiva usando una tecnica de barrido.
        % Se variaron los parametros FVP-FVP: $\alpha$ y $\beta$ realizando incrementos de a 0.2

        All runs used Euler method with a step of $\frac{1}{1000}$ $[s]$. For each scenario 
        and parameter combination, $10$ runs were made 
        and their metric results averaged and compared against each set of parameters. \\
        After numerous iterations of this process, this are the parameters which best fit all scenarios:

@@

            \begin{itemize}
                \item $K_{1}=80$ {[}N/m{]}, 
                \item $\gamma=10$ 
                \item $K_{2}=100$ {[}N/m{]} 
            \end{itemize}
            \item FVP-FVP interaction

            \begin{itemize}
                \item $\alpha=1000$ 
                \item $\beta=[0.4,\,0.6]$ (Uniform distribution) 
                \end{itemize}
            \item Pedestrian-FVP interaction

@@

    All results for each metric for the crossing scenario are presented on table \ref{table:metric-crossing}:

    \begin{table}[H]
        \begin{centering}
            \begin{tabular}{|c|c|c|}
                \hline 
                Metric  & {\scriptsize{FVPM}} & {\scriptsize{SFM ($\alpha=2000,\,\beta=0.08$)}}\tabularnewline
                \hline 
                \hline 
                $1$  & {\scriptsize{$15.0\pm4.3$}} & {\scriptsize{$23.2\pm6.4$}}\tabularnewline
                \hline 
                $2$  & {\scriptsize{$158.2\pm60.6$}} & {\scriptsize{$57.2\pm18.8$}}\tabularnewline
                \hline 
                $3$  & {\scriptsize{$1.26\pm0.02$ $[m/s]$}} & {\scriptsize{$1.28\pm0.01$ $[m/s]$}}\tabularnewline
                \hline 
                $4$  & {\scriptsize{$17.61\pm0.27$ $[s]$}} & {\scriptsize{$17.33\pm0.06$ $[s]$}}\tabularnewline
                \hline 
                $5$  & {\scriptsize{$22.61\pm0.07$ $[m]$}} & {\scriptsize{$22.52\pm0.07$ $[m]$}}\tabularnewline
                \hline 
                $6$  & {\scriptsize{$103.35\pm32.81$ $[rad]$}} & {\scriptsize{$193.06\pm20.44$ $[rad]$}}\tabularnewline
                \hline 
            \end{tabular}
        \par\end{centering}
        \caption{Results for crossing scenario \label{table:metric-crossing}}
    \end{table}

    Metrics for the crossing scenario don't present noticeable improvements over the SFM.
    The one improvement is the average turn angle, which can be recognized by looking
    and comparing the two models. The SFM presents a much more unnatural navigation than
    the FVPM.

    \begin{figure}[H]
        \begin{centering}
            \includegraphics[width=6cm]{pics/program/crossing-no-future} 
            \par
        \end{centering}
        \caption{\label{fig:crossing-no-future}Crossing visualization}
    \end{figure}

    \begin{figure}[H]
        \centering{} \includegraphics[width=6cm]{pics/program/crossing-future}
        \caption{\label{fig:crossing-future}Crossing visualization with visible future}
    \end{figure}
